---
title: "Time Series Models"
author: "Rizadh P H"
date: "21 September 2019"
output: html_document
---
Retail Sales time series Model to forecast sales for next two years sales per day

```{r}
# Load the forecasting package
library(fpp2)
library(ggplot2)

#Load the data
data <- read.csv("C:/Users/rizadh/Desktop/Retail_Sales_Adjusted.csv")

# Declare the sales column as time series 
Y <- ts(data[,2],start = c(1992,1), frequency = 12)
```

Time series plot of ts dataframe as an Initial step

```{r}
autoplot(Y) + 
  ggtitle("Time Plot")+
  ylab("Sales in Millions")
```











Major components of a time series are Cyclic, Noise, Trend and Seasonality. It is evident from the plot that it has Trend and Seasonality.

Taking the year on year difference can handle the trend 
```{r}
DY <- diff(Y)
```
Time plot for time series component after handling for trend
```{r}
autoplot(DY) + 
  ggtitle("Time Plot on difference")+
  ylab("Millions in sales")
```






No more trends now, so we can see if there is Seasonality now.

Seasonal subseries to understand the seasonality
```{r}
ggsubseriesplot(DY)+
  ggtitle("Sesonal Plot change in daily retail sales")+
  ylab("Millions in sales")
```







Yes, there is seasonality; one of the major instances is when we can see the sales is higher on december and low on January.


Time Series Models


Model 1 - Seasonal Naive method
```{r}
model1 <- snaive(DY)  
print(summary(model1))
summary(DY)
checkresiduals(model1)
```




Notable Result: Residual Standard Deviation  =  287.0647

Model 2 - Exponential smothing model
```{r}
model2 <- ets(Y)
print(summary(model2))
checkresiduals(model2)
```





Notable Result: Sigma(Residual Standard Deviation)  =  218.8133

Model 3 - ARIMA Model
```{r}
model3 <- auto.arima(Y,d=1,D=1,stepwise = FALSE,approximation = FALSE,trace = FALSE) 
print(summary(model3))
checkresiduals(model3)
```








Sigma^2 estimated as 39129 and hence Sigma (Residual Standard Deviation) = 197.81. Moreover, in the ACF plot majority data points lies within the range of statistical significance. Hence we can conclude that among these three model, ARIMA model is the better candidate for forecasting

Forecast - using ARIMA
```{r}
fcst <- forecast(model3,h=24)
```
Forecast series plot
```{r}
autoplot(fcst,include=120)
```






Forecast series results with High and low ranges
```{r}
print(summary(fcst))
```


